---
title: "TIRE cell input titration"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6, 
                      fig.path='./figures/cellInput_titrate/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(SingleCellExperiment)
library(tidyverse)
library(scuttle)
library(scater)
library(platetools)
library(patchwork)
library(here)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# TIRE cell input titration

## Aim

Investigate the dose response relationship of cell inputs to library size etc

## Samples
 
A serial dilution will be prepared on the day of the experiment.

* Human HEK293T
* Mouse NIH3T3
* No template control

[ELN entry](https://au-mynotebook.labarchives.com/share/Daniel_RnD_Projects/NDMyLjkwMDAwMDAwMDAwMDAzfDY4MTk3LzMzMy9UcmVlTm9kZS8yMjA0NzM2OTgzfDEwOTguODk5OTk5OTk5OTk5OQ==)

Standard TSO where UMI on TSO followed by spacer. Well barcode in line with read 1

## Data to read in

I sequenced these libraries with the configuration

* Read 1 70bp
* Read 2 52bp
* Index read 1 and 2 8bp each

## Read SCE and extract col data

```{r}
sce <- readRDS(here::here(
   "data/TIRE_technical/SCEs/cellInputs_basic.sce.rds"))

tb <- as_tibble(colData(sce))

# Keep only the 8 PCR samples as these performed best and had most reads
tb <- tb %>% 
  filter(PCR_cycles == 8)
```

# Check on plate map representation {.tabset}

Plot on plate map

## Samples

```{r plate_sample}

p1 <- platetools::raw_map(data=tb$Sample, 
                     well=tb$Well, 
                     plate=96)

p2 <- platetools::raw_map(data=log(tb$sum+1), 
                     well=tb$Well, 
                     plate=96)

p1 / p2
```

# All replicate dot plots by RNA amount {.tabset}

## Library size log

```{r all_libSize_log, fig.cap="Library size in deduplicated UMIs"}
plt1 <- ggplot(tb,
             aes(x = Amount+1, y= sum+1, colour = Sample)) + 
  geom_point(size=2) + 
  ylab("Library Size (UMIs)") + 
  xlab("Cell equivalents") +
  geom_smooth(method='lm', se = FALSE) +
  scale_y_continuous(trans='log10') +
  scale_x_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "l") +
  annotation_logticks(base =10, sides = "b") +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~Replicate, ncol=1)

plt1
```

## Library size natural scale

```{r all_libSize_nat, fig.cap="Library size in deduplicated UMIs"}
plt2 <- ggplot(tb,
             aes(x = Amount, y= sum, colour = Sample)) + 
  geom_point(size=2) + 
  ylab("Library Size (UMIs)") + 
  xlab("Cell equivalents") +  
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~Replicate, ncol=1)

plt2
```

## Detected genes

```{r all_detected_genes_p, fig.cap="Number of detected genes"}
plt3 <- ggplot(tb,
             aes(x = Amount, y= detected, colour = Sample)) + 
  geom_point(size=2) + 
  ylab("Genes detected") + 
  xlab("Cell equivalents") +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~Replicate, ncol=1)

plt3
```

# Show only 3rd replicate

Only replicate 1 HEK293T is problematic. However for the sake of clarity in publication I will 2nd replicate

```{r}
tb_orig <- tb

tb <- tb %>% 
  filter(Replicate == "Two")

tb$Amount <- tb$Amount / 1000
tb$Amount <- as.factor(tb$Amount)
```

## Library size log

```{r libSize_log, fig.cap="Library size in deduplicated UMIs"}
plt1 <- ggplot(tb,
             aes(x = Amount, y= sum+1, colour = Sample)) + 
  geom_point(size=2) + 
  ylab("Library Size (UMIs)") + 
  xlab("Cell equivalents (000s)") +
  geom_smooth(method='lm', se = FALSE) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "l") +
  scale_fill_brewer(palette = "Dark2")

plt1
```

## Library size natural scale

```{r libSize_nat, fig.cap="Library size in deduplicated UMIs"}
plt2 <- ggplot(tb,
             aes(x = Amount, y= sum, colour = Sample)) + 
  geom_point(size=2) + 
  ylab("Library Size (UMIs)") + 
  xlab("Cell equivalents (000s)") +  
  scale_fill_brewer(palette = "Dark2")

plt2
```

## Detected genes

```{r detected_genes_p, fig.cap="Number of detected genes"}
plt3 <- ggplot(tb,
             aes(x = Amount, y= detected, colour = Sample)) + 
  geom_point(size=2) + 
  geom_smooth(method='lm', se = FALSE) +
  ylab("Genes detected") + ylim(0, NA) +
  xlab("Cell equivalents (000s)") +
  scale_fill_brewer(palette = "Dark2")

plt3
```

# Human and mouse mapping {.tabset}

## Barnyard natural scale

```{r barnyard}
p1 <- ggplot(tb, aes(x=subsets_Mouse_sum, y=subsets_Human_sum, colour=Sample)) + 
    geom_point(size=2) +
  xlab("Mouse library size") + 
  ylab("Human library size") +
  ylim(0,2e6) + xlim(0,2e6) +
  scale_colour_brewer(palette = "Dark2")

p1
```

## Barnyard log scale

* Looks deceptive and not how anyone else shows this.  
* Replicate 2 nd 3 look very clean even in log representation

```{r species_counts_log}
p2 <- ggplot(tb, aes(y=subsets_Human_sum, x=subsets_Mouse_sum, colour=Sample)) + 
    geom_point(size=2) +
  xlab("Mouse library size") + 
  ylab("Human library size") +
  scale_y_continuous(trans='log10') + scale_x_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "bl") +
  scale_colour_brewer(palette = "Dark2")

p2
```

# Conclusion

* Replicate 1 looks compromised in terms of contamination especially one of the wells
* Replicate 1 HEK293T has contamination issues
* Otherwise all other wells within replicate 2 and replicate 3 have worked.

### Session info
```{r}
sessionInfo()
```