---
title: "TIRE vs Prime-seq barnyard"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6, 
                      fig.path='./figures/barnyard/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(SingleCellExperiment)
library(tidyverse)
library(scuttle)
library(scater)
library(platetools)
library(patchwork)
library(here)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# TIRE vs Prime-seq barnyard

## Aim

Investigate the difference between species separation in barnyard plots

## Data to read in

I read in experiments where I spiked my own human and mouse cell lines; HEK293T and NIH3T3.

TIRE-seq / TurboCapture-seq [mouse DC expt][https://au-mynotebook.labarchives.com/share/Daniel_RnD_Projects/NDUzLjd8NjgxOTcvMzQ5L1RyZWVOb2RlLzEzMDk5NDU5MDZ8MTE1MS43]

* Prime-seq/ Minibulk v2. [Splicing experiment 1](https://au-mynotebook.labarchives.com/share/Daniel_RnD_Projects/NDUyLjQwMDAwMDAwMDAwMDAzfDY4MTk3LzM0OC9UcmVlTm9kZS8xNjc4OTU2MTQwfDExNDguMzk5OTk5OTk5OTk5OQ==)

## Read SCE and extract col data

```{r}
tire <- readRDS(here::here(
  "data/TIRE_dendritic_mouse/SCEs/", "DCs_Mixed_basic.sce.rds"))

tb_tire <- as_tibble(colData(tire))
tb_tire$Protocol <- "TIRE-seq"

# Keep only the 8 PCR samples as these performed best and had most reads
tb_tire <- tb_tire %>% 
  filter(Researcher == "DB")

tb_tire$Species <- recode(tb_tire$Species,
                         "None" = "NTC")
```

Prime-seq/ minibulk v2. In this experiment replicate 3 had the Janjic et al., style of RT primer removal; 2 bead cleanups + exo treatment.  
The first 2 replicates inherit the RT removal of our old protocols. 1 exo treatment and 1 bead cleanup.

```{r}
mb2 <- readRDS("/stornext/Home/data/allstaff/b/brown.d/Projects/G000233_DRAC-Seq/data/MB2_Benchmarking/SCEs/benchmarking_expt1_basic.sce.rds")

mb_tb <- as_tibble(colData(mb2))
mb_tb$Protocol <- "Modified Prime-seq"

# Remove first replicate leaving only 1 replicate each of different primer removal stratgeises
mb_tb <- mb_tb %>% 
  filter(Researcher == "DB") %>% 
  filter(Plate_replicate == "Three") %>% 
  filter(Biological_replicate == 1)
```

Combine the tibbles together for subsequent visualisation.

```{r}
keep_cols <- intersect(
  colnames(mb_tb),
  colnames(tb_tire)
)

tb <- rbind(
  mb_tb[,keep_cols],
  tb_tire[,keep_cols]
)
```

# Human and mouse mapping {.tabset}

Add a CPM column to make the scale easier to see.

```{r}
tb <- tb %>% 
  group_by(Protocol) %>% 
  mutate(
    mouse_CPM =  subsets_Mouse_sum / sum(sum) * 1e6,
    human_CPM =  subsets_Human_sum / sum(sum) * 1e6
  )
```

## Barnyard natural scale

```{r barnyard}
p1 <- ggplot(tb, aes(x=subsets_Mouse_sum, y=subsets_Human_sum, colour=Protocol, shape=Species)) + 
    geom_point(size=2) +
  xlab("Mouse library size") + 
  ylab("Human library size") +
  ylim(0,2e6) + xlim(0,2e6) +
  scale_colour_brewer(palette = "Dark2")

p1
```

## Barnyard cpm scale

This is not very useful

```{r barnyard_cpm}
p3 <- ggplot(tb, aes(x=mouse_CPM, y=human_CPM, colour=Protocol, shape=Species)) + 
    geom_point(size=2) +
  xlab("Mouse library size") + 
  ylab("Human library size") +
  ylim(0,3e5) + xlim(0,3e5) +
  scale_colour_brewer(palette = "Dark2")

p3
```

## Barnyard log scale

Probably the fairest looking comparison to have everything looking similar

```{r species_counts_log}
p2 <- ggplot(tb, aes(y=subsets_Human_sum, x=subsets_Mouse_sum, colour=Protocol, shape=Species)) + 
    geom_point(size=2) +
  xlab("Mouse library size") + 
  ylab("Human library size") +
  scale_y_continuous(trans='log10', limits = c(1,2e6)) + scale_x_continuous(trans='log10', limits = c(1,2e6)) +
  annotation_logticks(base = 10, sides = "bl") +
  scale_colour_brewer(palette = "Dark2")

p2
```

## Violin percentages

Plotting the percent mapped also looks like a reasonable representation.

As there are fewer mouse samples in TIRE-Seq one would expect to see more bleed through of human into mousse which is not the case.

```{r}
percent <- tb %>% 
  filter(Species != "NTC") %>% 
  filter(Protocol != "Minimal primer removal") %>% 
  select(Species, subsets_Human_percent, subsets_Mouse_percent, Protocol) %>% 
  pivot_longer(cols = c(subsets_Human_percent, subsets_Mouse_percent), names_to = "Subset", values_to = "Percent")

percent$Subset <- recode(percent$Subset,
                         subsets_Human_percent = "Human",
                         subsets_Mouse_percent = "Mouse"
                         )
```

Produce the visualisation

```{r violin_percent}
filtered_tibble <- percent %>%
  filter((Species == "Mouse" & Subset == "Human") |
         (Species == "Human" & Subset == "Mouse"))

filtered_tibble$Subset <- recode(filtered_tibble$Subset,
                                 "Human" = "Mouse sample -> \nMapped human",
                                 "Mouse" = "Human sample -> \nMapped mouse")

p5 <- ggplot(filtered_tibble, aes(y=Percent, x=Subset, colour=Protocol)) + 
    geom_jitter(width = 0.15, alpha=0.75) +
  xlab("") + 
  ylab("Percent mapped") + ylim(0,15) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
  scale_colour_brewer(palette = "Dark2")

p5
```

Look at the number of samples in each category

```{r}
filtered_tibble %>% 
  group_by(Subset, Protocol) %>% 
  count()
```



### Session info
```{r}
sessionInfo()
```