---
title: "UHRR protocol comparisons"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='./figures/downsampling/UHRR/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(tidyverse)
library(scuttle)
library(here)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# A protocol comparison of Minibulk and TurboCapture on UHRR

I reproduced the experiment from the Prime-Seq experiment where universal human reference RNA (UHRR) was mixed with ERCC spike-in RNAs with the protocols I implemented:

* Minibulk v2 (modified Prime-Seq)
* TIRE-Seq (Turbocapture Integrated RNA Expression)

## Read data

```{r}
tb <- read_csv(here::here(
  "data/TIRE_UHRR/Matrices/downsampling_stats.csv"
))
tb
```

There are not enough samples with more than 10M reads so filter at this point.

```{r}
tb <- tb %>% 
  filter(reads <= 10000000)

tb$Protocol <- recode(tb$Protocol,
                      "Minibulk v2" = "Modified Prime-seq",
                      "Janjic et al" = "Janjic Prime-seq"
)
```

Read in the TruSeq sample. Reviewer has requested the addition of TruSeq data. This was from the SEQC consortium sample A.

```{r}
truseq <- read_csv("/home/users/allstaff/brown.d/Projects/G000278_TurboCapSeq/data/TruSeq_UHRR_public/Matrices/truSeq_UHRR_downsampling_stats.csv")
truseq$Protocol <- "TruSeq"

truseq$reads <- str_split(truseq$id, pattern = "_", simplify = T)[,2]
truseq$reads <- as.numeric(truseq$reads)

truseq <- truseq %>% 
  filter(reads <= 2.5e6)

keep_col <- intersect(
  colnames(tb),
  colnames(truseq)
)

tru_tb <- rbind(
  tb[,keep_col], truseq[,keep_col]
)
```

## Visualise the results {.tabset}

In both visualizations TIRE has better efficiency but saturates quicker

### Genes detected

TIRE-Seq saturates quicker while minibulk v2 keeps increasing

```{r downsample_gene, fig.cap="Error bars are median absolute deviation"}
plt1 <- ggplot(data=tb, 
               aes(x=reads, y=med_gene, group=Protocol, color=Protocol)) +
  geom_line(size=1.25) + geom_point(size=2) +
  geom_errorbar(aes(ymin=med_gene-mad_gene, ymax=med_gene+mad_gene), width=.1, size=1.25) +
  scale_colour_brewer(type="Qualitative", palette = "Dark2") +
  ggtitle("Purified RNA") +
  xlab("Reads per sample") + ylab("Median genes detected")

plt1
```

Visualise the TruSeq data on the saturation plot.

```{r downsample_gene_truseq, fig.cap="Error bars are median absolute deviation"}
plt1 <- ggplot(data=tru_tb, 
               aes(x=reads, y=med_gene, group=Protocol, color=Protocol)) +
  geom_line(size=1.25) + geom_point(size=2) +
  geom_errorbar(aes(ymin=med_gene-mad_gene, ymax=med_gene+mad_gene), width=.1, size=1.25) +
  scale_colour_brewer(type="Qualitative", palette = "Dark2") +
  ggtitle("Purified RNA") +
  xlab("Reads per sample") + ylab("Median genes detected")

plt1
```

### UMIs detected

It is more stark here that minibulk keeps increasing while TIRE-Seq is more efficient but saturates more quickly.  
I checked on the the highly expressed genes and this is not a UMI saturation issue.

```{r downsample_umi, fig.cap="Error bars are median absolute deviation"}
plt2 <- ggplot(data=tb, 
               aes(x=reads, y=med_umi, group=Protocol, color=Protocol)) +
  geom_line(size=1.25) + geom_point(size=2) +
  geom_errorbar(aes(ymin=med_gene-mad_gene, ymax=med_gene+mad_gene), width=.1, size=1.25) +
  scale_colour_brewer(type="Qualitative", palette = "Dark2") +
  xlab("Reads per sample") + ylab("Median UMIs detected")

plt2
```

### Number of samples at depth

Prime-Seq and TurboCap stands out from the minibulk implementations.

```{r downsample_wells, fig.cap="Error bars are median absolute deviation"}
plt3 <- ggplot(data=tb, 
               aes(x=reads, y=cells, group=Protocol, color=Protocol)) +
    geom_line() + geom_point() +
  scale_colour_brewer(type="Qualitative", palette = "Dark2") +
  xlab("Reads per sample") + ylab("Number of samples")

plt3
```

### Genes zoom 1 million

```{r uhrr_genes_2M}
plt1 + xlim(0,1e6) + ylim(0,25000)
```

### UMIs zoom 1 million

```{r uhrr_umi_2M}
plt2 + xlim(1,1e6) + ylim(0,1e6)
```

# Conclusions

* TIE-Seq saturates quicker while minibulk v2 keeps increasing
* It is more stark here that minibulk keeps increasing while TIRE-Seq is more efficient but saturates more quickly.  
* I checked on the the highly expressed genes and this is not a UMI saturation issue.

## Next steps

Count the reads with Gencode annotation next time and perform analysis again

### Session info
```{r}
sessionInfo()
```
