---
title: "Brain cancer visualise quality control"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6, 
                      fig.path='./figures/qc_metrics/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(SingleCellExperiment)
library(tidyverse)
library(scuttle)
library(scater)
library(platetools)
library(patchwork)
library(here)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# TurboCapture-Seq Brain Cancer neurospheres

## Recap

These are the first experiments where I use version 2 chemistry. Read 2 is now the barcode and UMI Read 1 is the cDNA. This was done to avoid reading through the TSO on Ilumina read 1

Test version 2 protocol with a biological appliation from Zac Moore of Brain Cancer lab [https://au-mynotebook.labarchives.com/share/Piper_Research_Project/MzEuMjAwMDAwMDAwMDAwMDAzfDE3MTc2NS8yNC9UcmVlTm9kZS8zMzczNDM0NjE1fDc5LjE5OTk5OTk5OTk5OTk5](ELN link)

## Read SCE and extract col data

This is the SCE object from single end umi

```{r}
sce <- readRDS(here::here(
  "data/TIRE_brain_human/SCEs/", "brainCancer_basic.sce.rds"))

tb <- as_tibble(colData(sce))

tb$Species[is.na(tb$Cell_Line)] <- "None"
tb$Cell_Line[is.na(tb$Cell_Line)] <- "None"
```

# Check orientation on plate map representation {.tabset}

Orientation is correct when looking at library size and mouse counts.

## Samples

```{r plate_sample, fig.width=12,fig.height=6}
p1 <- platetools::raw_map(data=tb$Cell_Line, 
                     well=tb$Well, 
                     plate=96) + scale_fill_Publication()

p2 <- platetools::raw_map(data=log(tb$sum+1), 
                     well=tb$Well, 
                     plate=96) +
  ggtitle("Library Size")


p3 <- platetools::raw_map(data=tb$Day_Exposure, 
                     well=tb$Well, 
                     plate=96) + 
  ggtitle("Days exposure to drug")

p4 <- platetools::raw_map(data=tb$Storage, 
                     well=tb$Well, 
                     plate=96) + 
  ggtitle("Storage conditions") + scale_fill_Publication()

p1 + p2 + p3 + p4
```

## Conditions

```{r plate_conditions, fig.width=12,fig.height=6}
tb_fact <- tb 
tb_fact$Day_Exposure <- as.factor(tb$Day_Exposure)
tb_fact$Dose_M <- as.factor(tb$Dose_M)

p1 <- platetools::raw_map(data=tb_fact$Species, 
                     well=tb$Well, 
                     plate=96) + 
  scale_fill_discrete("Species")

p2 <- platetools::raw_map(data=tb_fact$Day_Exposure, 
                     well=tb$Well, 
                     plate=96) +
  scale_fill_discrete("Timepoint (days)")

p3 <- platetools::raw_map(data=tb_fact$Dose_M, 
                     well=tb$Well, 
                     plate=96) + 
  guides(fill=guide_legend(ncol=2)) +
  scale_fill_discrete("Dose (M)")

p4 <- platetools::raw_map(data=tb_fact$Drug, 
                     well=tb$Well, 
                     plate=96) + 
  scale_fill_discrete("Drug")

p1 + p4 + p3 + p2
```

# Quality control metrics by dose {.tabset}

The metrics go as one would expect with dose and time treated

## Library size

```{r treat_lib}
tb$Day_Exposure <- as_factor(tb$Day_Exposure)

plt1 <- ggplot(tb,
             aes(x = Dose_M, y= sum, colour=Day_Exposure)) + 
  geom_point() +
  ylab("Library Size (UMIs)") + 
  xlab("TMZ (M)") +
  scale_y_continuous(trans='log10') + scale_x_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "bl") +
  scale_colour_brewer(palette = "Dark2")

plt1
```

## Genes detected

```{r treat_genes}
plt2 <- ggplot(tb,
             aes(x = Dose_M, y= detected, colour=Day_Exposure)) + 
  geom_point() +
  xlab("TMZ (M)") +
  ylab("Genes detected") + 
  scale_x_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "b") +
  scale_colour_brewer(palette = "Dark2")

plt2
```

## Mitochondrial percent

```{r treat_mito}
plt2 <- ggplot(tb,
             aes(x = Dose_M, y= subsets_Mito_percent, colour=Day_Exposure)) + 
  geom_point() +
  xlab("TMZ (M)") +
  ylab("Mitochondrial percent") + 
  scale_x_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "b") +
  scale_colour_brewer(palette = "Dark2")

plt2
```

# Investigate storage conditions {.tabset}

Evidence that storage at 4C is insufficient to freeze the biological effect of temozolomide on the library size. In general metrics are poorer for storage at 4C.

```{r}
storage <- tb %>% 
  filter(Day_Exposure == 7) %>% 
  filter(Dose_M != 0.0000001) %>% 
  filter(Dose_M != 0.000001) %>% 
  filter(Dose_M != 0.00000316) %>% 
  filter(Dose_M != 0.000000316) %>% 
  filter(Dose_M != 0.0000316)
  
storage %>% 
  dplyr::count(Dose_M,Storage)
```

## Library size

```{r storage_lib}
plt1 <- ggplot(storage,
             aes(x = Storage, y= sum, colour=Drug)) + 
  geom_jitter(width = 0.1) + 
  ylab("Library Size (UMIs)") + 
  xlab("Storage") +
  scale_y_continuous(trans='log10') +
  annotation_logticks(base = 10, sides = "l") +
  scale_colour_brewer(palette = "Dark2")

plt1
```

## Genes detected

```{r storage_genes}
plt2 <- ggplot(storage,
             aes(x = Storage, y= detected, colour=Drug)) + 
  geom_jitter(width = 0.1) + 
  ylab("Genes detected") + 
  xlab("Storage") +
  scale_colour_brewer(palette = "Dark2")

plt2
```

## Mitochondrial percent

Further evidence that 4C storage does not inactivate Temozolomide effects.

```{r storage_mito}
plt3 <- ggplot(storage,
             aes(x = Storage, y= subsets_Mito_percent, colour=Drug)) + 
  geom_jitter(width = 0.1) + 
  ylab("Mitochondrial %") + ylim(0,15) +
  xlab("Storage") +
  scale_colour_brewer(palette = "Dark2")

plt3
```
 
### Session info
```{r}
sessionInfo()
```